Design and Perceptual Evaluation of a urban sound scenes corpus with controlled acoustic properties

### 2017-06-19 articleBoston.tex
add boston-presentation

\documentclass[a4,11pt,twocolumn]{article}

\input{packageBoston}

\title{Design of a urban sound scenes corpuses with controlled acoustic properties}     %% \title est une macro, entre { } figure son premier argument
\author{
  GLOAGUEN, Jean-R\'emy\      \texttt{first1.last1@xxxxx.com}
  \and
  CAN, Arnaud\      \texttt{first2.last2@xxxxx.com}
  \and
  LAGRANGE, Mathieu\      \texttt{first1.last1@xxxxx.com}
  \and
  PETIOT, Jean-Fran\c cois\      \texttt{first1.last1@xxxxx.com}
}

@@

\maketitle

\section*{Abstract}

Monitoring the acoustic urban environment is beneficial for improving the life of citizens. To do so, the research community in acoustics can befenit from corpora where the acoustic contribution of sources of disturbance such as road traffic is precisely known.

In this paper, a set of urban sound mixtures are simulated using an open source tool \textit{simScene}. Their caracteristics (type of sources, density of events, etc.) are infered from the listening of reference audio scenes recorded in Paris, France. The main advantage of the simulated versions is that several quantities, such as the acoustic level of each source is precisely known, a useful feature for evaluated automated analysis tools.

In order to quantifiy the level of realism achieved by the use of the simulation tool a perceptive test is conducted with 50 subjects. The result show that no significative differences can be made between the simulated and the recorded scenes.

\section*{Introduction}

The urban sound environment is studied both from the perception \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and from the physical point of view \cite{can_describing_2015} \cite{raimbault_ambient_2003}. These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recording made in city \cite{botteldooren_temporal_2006} or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}. In the first case, if the urban sound environment have an undeniable ecology validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted. Then the use of a simulation tool to create specific urban sound environments is a useful way to control the degree of contribution of each class of acoustical sources. It allows us to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to evaluate the performance of some classification or detection tools \cite{giannoulis_detection_2013}.

A critical question is then to obtain scenes that are realistic enough, thus to be considered as similar to some reference recordings. To do so, various methods have been proposed. The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamic of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}. If the first renderings are interesting, this method remains complex to implement.
A more straitghtforward approach composes sound mixtures from isolated sound mixed together. The idea is to consider the urban sound environment as the sum of acoustic events, i.e. events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene. The difficulty here is to have a representative database of isolated sounds with a sufficient quality not to compomise the resulting sound quality and to sequence the acoustic events wisely.

The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical event directly in real recordings and by manipulating them to re-use them in new mixtures. If the tool has a sufficient number of parameters to control the sound event, their method is limited by the extraction phase: the overlapping between the acoustics events deteriorates the sounds and then the final rendering. In more, the manipulation (duration, frequency range) can create artifacts that can be a limitation for a realistic use. In a similar way, the tool proposed by Davis and Bruce \cite{bruce_development_2009} enables to compose sound mixtures with an isolated sound database composed of recordings made specially for their application. The creation of the urban sound mixtures is made possible with the perceptual evaluation of a panel of listeners of the different sound environments. This method enables to determine which sounds are the most representative or the most linked to an urban sound environment from a perspective point of view.

However, as our objective here is to get simulated scenes that are as realistic as possible, this method can not be considered because if this method proposes a perspective validity it can not provide an ecological one. Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some high level sequencing parameters extracted from the recordings to tune an audio simulation tool. The realism of the simulated scenes is then evaluated using a perspective test.

The remaining of the paper is organized as follows, Section \ref{sec:study} focuses on the study of recordings made in Paris. Section \ref{sec:simulation} deals with the simulation tool and Section \ref{sec:test} summarizes the result obtained in the perspective test.

\section{Study of realistic scenes} \label{sec:study}
Real urban noise recordings are listened in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level. The recordings considered in this study have been made on the \nth{13} district of Paris (France) as part of the GRAFIC project \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments. The walk was 2.1 km long, consisting of 19 stop points of 3 to 5 minutes of listening each (see figure \ref{fig:soundwalk}). It was roamed on two days (03/23/2015 and 03/30/2015) twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW). The acquisition system used was a ASA SENS and was carried on the backpack of the operator which was walking with the participants. In the end, 74 audio pulse code modulation recordings sampled at 44,1 kHz are available and serve as a base for the study. More details can be found on  \cite{aumond_modelling_2017}. \\

@@

Each audio is fully annotated in terms of sound classes that are be present in a urban area, their recurrence and is labeled using the following types of sound environments (\textit{park}, \textit{quiet street} , \textit{noisy street}, \textit{very noisy street}) as proposed by \cite{rychtarikova_soundscape_2013} or \cite{can_describing_2015} (table~\ref{tab:classificationScene}).

@@

\begin{tabular}{|c|c|*{19}{l|}}
\hline
\multicolumn{1}{|l|}{\textbf{day}} & \textbf{journey}   & 1                        & 2                        & 3                        & 4                        & 5                        & 6                        & 7                        & 8                        & 9                        & 10                       & 11                       & 12                       & 13                       & 14                       & 15                       & 16                       & 17                       & 18                       & 19                       \\ \hline
\multirow{2}{*}{\textbf{1}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
\cline{2-20}
 & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} &                          & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} &  \\
 \hline
\multirow{2}{*}{\textbf{2}} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
\cline{2-20}
 & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{5AB25A} \\ \hline
\end{tabular}

\vspace{0.5cm}

@@

Most of the scene of the recordings belong to \textit{quiet street} and \textit{noisy street}. 2 scenes are not available : a noisy cleaning truck is present throughout the \textit{1-WE-3} heavily polluting the sound scene and the \textit{1-WE-19} record is too short to be considered in this study. From the annotation, each sound ambiance is characterized with the sound classes presents in it. The most common sound classes are displayed in Table \ref{tab:obsScene}.\\

\begin{table*}[!h]
\centering
\begin{tabular}{L{2.5cm} | C{2.5cm} | c | C{3cm} | C{2.5cm}}
\centering \textbf{Sound environment} & \textbf{Sound level (dB)} & \textbf{Background                                                       } &\textbf{Event} & \textbf{number events/min} \\ \hline
Park              & 69.0        &
\begin{tabular}[c]{@{}l@{}}voice, \\ bird's whistles\end{tabular} &
\begin{tabular}[c]{@{}l@{}}road traffic \\ voices\\ bird's whistle\\ street noise\\ foot step\end{tabular}                                         &
\begin{tabular}[c]{@{}l@{}}0.5\\ 0.5\\ 0.5\\ 0.5\\ 0.3\end{tabular}                       \\\hline
Quiet street      & 70.2        &
\begin{tabular}[c]{@{}l@{}}road traffic\\ bird\end{tabular}       &
\begin{tabular}[c]{@{}l@{}}road traffic \\ voices\\ street noises\\ foot step\\ bird\\ \begin{tabular}[c]{@{}l@{}}construction \\ \hfill site noise \end{tabular}\\ door house\\ door car\end{tabular} &
\begin{tabular}[c]{@{}l@{}}1.0\\ 0.7\\ 0.7\\ 0.5\\ 0.2\\ \begin{tabular}[c]{@{}l@{}} 0.1\\~\end{tabular}\\  0.2\\ 0.2\end{tabular}
\\ \hline
Noisy street      & 73.5        & road traffic                                                      & \begin{tabular}[c]{@{}l@{}}traffic\\ foot step\\ voice\\ street noise\\ bell\\ bird\\ car horn\\ car's door\\ siren\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}9.0 \\ 0.5\\ 0.6\\ 0.4\\ 0.1\\ 0.2\\ 0.3\\ 0.2\\ 0.1\end{tabular} \\ \hline
Very noisy street & 76.0        & road traffic                                                      & \begin{tabular}[c]{@{}l@{}}traffic\\ voice\\ siren\\ car horn\\ bird\\ foot step\\ car's door\\ street noise\end{tabular}                          & \begin{tabular}[c]{@{}l@{}}40\\ 0.3\\ 0.2\\ 0.3\\ 0.2\\ 0.3\\ 0.2\\ 0.3\end{tabular}
\end{tabular}
\caption{Sound level and classes presents depending on the sound environment}
\label{tab:obsScene}
\end{table*}

%\begin{table*}[t]
%\centering
%
%\begin{tabular}{p{3cm}p{2cm}p{10.5cm}}
%\hline
%\multicolumn{1}{c}{\textbf{Ambience}} &\textbf{\begin{tabular}[c]{@{}c@{}}level\\ (dB SPL)\end{tabular}} & \multicolumn{1}{c}{\textbf{Description}} \\ \hline
%park & \centering 69,0 & predominance of children's and adults' voices and bird whistles as sound events and backgrounds, the road traffic is low almost non-existent ($\approx$ 1 car/min) \\
%\hline
%quiet street & \centering 70,2 & The background noise is caused by a low road traffic with few car passages ($\approx$ 2 car/min), the birds are present and audible, the voice is composed of children's and adult's voices. The effect of reverberation on the street, due to low background noise, is present.\\
%\hline
%animated street & \centering 73,5 & traffic noise  is louder and the unique background, the car passages density is higher ($\approx$ 10 car/min), the human activity is more present and diversify (voice, step, suitcase, construction site noise \dots), the birds disappear. \\
%\hline
%very noisy street & \centering 76,0 & the background noise very high composed of quasi-constant traffic flow ($\sup$ 40 car/min), the human activity are less presents in a favor of voice hubbub.\\
%\hline
%
%\end{tabular}
%\caption{Qualitative observations on the urban sound environment}
%\label{tab:obsScene}
%\end{table*}

The road traffic, the voice and the bird components are the most linked to the sound atmosphere: with exclusively road traffic, the sound environment is more noisy than in the \textit{park} and \textit{quiet street} where the voice and the bird's whistles are predominant. Furthermore, multiple sounds classes can be heard whatever the urban sound environments : dog barks, church bell ringing, car horn, sirens \dots with a density which can be very low (< 0.1/min). Finally, a lot of brief mechanics sounds with unknown origin can be listened in almost all the audio. Consequently, all these indefinite sounds are annotated in one unique class sound, \textit{street noise}. From theses information (density, level, sound classes present in each ambience), it is now possible to compose urban sound mixtures.

\section{Simulation of realistic simulated urban scenes} \label{sec:simulation}
\subsection{Presentation of the web-simulator \textit{simScene}}
\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound mixtures in a wav format by superposing audio which comes from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds)
that are considered as salient belong to the \textit{event} category and all the sound that are of long duration and whose acoustic properties do not vary with respect to time belong to the \textit{background} category. In each category, the sounds are then grouped in sound classes (bird, car, foot steps \dots). The software enables the user to control for each sound classes some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots. Each parameters are completed with a standard deviation optionally bringing some random behavior between the scenes. \\

\begin{figure}[h]
\centering
\includegraphics[width = \linewidth]{./pictures/spectrogramme_abstract_sceneSimpleKlaxonFixe_EN.pdf}
\caption{Example of a scene composed of three sound classes}
\label{fig:exampleSimScene}
\end{figure}

The sound mixtures can be created following two modes: the \textit{abstract} mode allows us to create mixtures from the specified parameters while the \textit{replicate} mode reproduces an existing scene by following as much as possible its annotation file. With the global sound mixture, an audio can also be created for each sound class present in the scene which allows to know their exact contribution in the scene. %In parallel, \textit{simScene} generates 3 image files representing the temporal evolution, the spectrogram and a piano Roll representation to visualize each sound classes distribution and also two text files summarizing the temporal location of each sample and all the parameter used. \\

\subsection{Creation of the urban sound database}
Based on the annotations of the real scenes, a sound database is built using the \textit{replicate} mode with sounds found online on the \textit{freesound} project and with the urbanSound8k database\cite{salamon_dataset_nodate}. This database consists in more than 8000 files with a 4 seconds or less duration found on the web site freesound.org too, and classified in 10 sound classes : ventilation, car horn, children playing, dog barking, bell, engine idling, gunshot, jackhammer, siren and street music. % ML Pas clair, est-ce utile ? All the audio have been sorted and those who presented the best signal noise ratio are selected in the sound classes identified in the real records have been selected.\\

In addition, as road traffic is a prime audio source in an urban sound environment, it seemed interesting to have, in the database, a part composed of well-recorded car audio. As a result, passages of 4 different cars (Renault Scenic, Clio, Megane and Dacia Sandero) were recorded on the Ifsttar-Nantes runway at different speeds and gear ratios for steady, acceleration and braking phases (table~\ref{tab:resum_car_audio}). In all, 103 car passages have been recorded. \\

>>>>

\documentclass[11pt,letter]{article}

\usepackage{poma_style}

\input{packageStyle}

\newcommand{\eps}{\varepsilon}

@@

%\section*{Abstract}
%
%Monitoring the acoustic urban environment is beneficial for improving the life of citizens. To do so, the research community in acoustics can benefit from corpora where the acoustic contribution of sources of disturbance such as road traffic is precisely known.
%
%In this paper, a set of urban sound mixtures are simulated using an open source tool \textit{simScene}. Their characteristics (type of sources, density of events, etc.) are inferred from the listening of reference audio scenes recorded in Paris, France. The main advantage of the simulated versions is that several quantities (such as the acoustic level of each source, a useful feature automated analysis tools) is precisely known.
%
%In order to quantify the level of realism achieved by the use of the simulation tool a perceptive test is conducted with 50 subjects. The results show that no significant differences can be made between the simulated and the recorded scenes.

\section{Introduction}

The urban sound environment is studied both from the perceptual \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and from the physical point of view \cite{can_describing_2015} \cite{raimbault_ambient_2003}. These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recordings made in city \cite{botteldooren_temporal_2006}, or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}. In the first case, since it has an undeniable ecological validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted. Then the use of a simulation tool to create specific urban sound environments is a useful way to control the degree of contribution of each class of acoustical sources. It allows us to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to evaluate performances of some classification or detection tools \cite{giannoulis_detection_2013}.

A critical question is then to obtain scenes that are realistic enough to be considered simulated scenes as similar to some reference recordings. To do so, various methods have been proposed. The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamics of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}. If the first renderings are interesting, this method remains complex to implement.
A more straightforward approach composes sound mixtures from isolated sounds mixed together. The idea is to consider the urban sound environment as the sum of acoustic events, i.e. events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene. The difficulty here is to have a representative database of isolated sounds with a sufficient quality to not compromise the resulting sound quality and to sequence the acoustic events wisely.

The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical events directly in real recordings and by manipulating them to re-use them in new mixtures. Since the tool has a sufficient number of parameters to control the sound event, their method is limited by the extraction phase: the overlapping between the acoustic events deteriorates the sounds and then the final rendering. Furthermore, the manipulations (duration, frequency range) can create artifacts that can deacrease the realism. In a similar way, the tool proposed by Davis and Bruce \cite{bruce_development_2009} enables to compose sound mixtures with an isolated sound database composed of recordings made specially for their application. The creation of the urban sound mixtures is made possible with the perceptual evaluation of a panel of listeners of the different sound environments. This method enables to determine which sounds are the most representative or the most linked to an urban sound environment from a perceptual point of view.

However, as our objective here is to get simulated scenes that are as realistic as possible, this method can not be considered because if this method proposes a perceptual validity, it can not provide an ecological one. Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some high level sequencing parameters extracted from the recordings to tune an audio simulation tool. The realism of the simulated scenes is then evaluated using a perceptual test.

The remaining of the paper is organized as follows: Section \ref{sec:study} focuses on the study of recordings made in Paris, Section \ref{sec:simulation} deals with the simulation tool and Section \ref{sec:test} summarizes the results obtained in the perspective test.

\section{Study of real scenes} \label{sec:study}
Real urban noise recordings are listened in details to determine the composition of typical urban sound environments in terms of sound sources, presence and level. The recordings considered in this study have been made on the \nth{13} district of Paris (France) as part of the GRAFIC project \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments.\\

@@

The walk was 2.1 km long, consisting of 19 stop points with a recording duration between 3 to 5 minutes (see figure \ref{fig:soundwalk}). It was roamed on two days (03/23/2015 and 03/30/2015), twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW). The acquisition system used was a ASAsense and was carried on the backpack of the operator which was walking with the participants. In the end, 76 (4 $\times$ 19) audio files (sampled at 44,1 kHz) were available and serve as a basis for the study. More details can be found on \cite{aumond_modelling_2017}. Each file is fully annotated in terms of sound classes that are present in the urban area and their recurrence. It is labeled using the following types of sound environments (\textit{park}, \textit{quiet street}, \textit{noisy street}, \textit{very noisy street}) as proposed by \cite{rychtarikova_soundscape_2013} or \cite{can_describing_2015} (Table~\ref{tab:classificationScene}).

@@

\begin{tabular}{|c|c|*{19}{l|}}
\hline
\multicolumn{1}{|l|}{\textbf{day}} & \textbf{journey}   & 1                        & 2                        & 3                        & 4                        & 5                        & 6                        & 7                        & 8                        & 9                        & 10                       & 11                       & 12                       & 13                       & 14                       & 15                       & 16                       & 17                       & 18                       & 19                       \\ \hline
\textbf{1} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
\hline
\textbf{1} & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} &                          & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} &  \\
 \hline
\textbf{2} & \textbf{EW} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} \\
\hline
\textbf{2} & \textbf{WE} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{5AB25A} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{F56B00} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{FFCB2F} & \cellcolor[HTML]{9A0000} & \cellcolor[HTML]{5AB25A} \\ \hline
\end{tabular}
%
\vspace{0.5cm}

@@

Most of the scenes of the recordings belong to \textit{quiet street} and \textit{noisy street}. 2 scenes are removed form the database: a noisy cleaning truck is present inside the \textit{1-WE-3} heavily polluting the sound scene and the \textit{1-WE-19} record is too short to be considered in this study. From the annotations, each sound environment is characterized by the sound classes present in the recordings. The most recurrent sound classes are presented in Table \ref{tab:obsScene}.\\

\begin{table*}[!h]
\centering
\begin{tabular}{L{2.5cm} | C{2.5cm} | c | C{3cm} | C{2.5cm}}
\centering \textbf{Sound environment} & \textbf{Sound level (dB)} & \textbf{Background                                                       } &\textbf{Event} & \textbf{number events/min} \\ \hline
Park              & 69.0        &
\begin{tabular}[c]{@{}l@{}}voice, \\ bird's whistles\end{tabular} &
\begin{tabular}[c]{@{}l@{}}road traffic \\ voices\\ bird's whistle\\ street noise\\ foot step\end{tabular}                                         &
\begin{tabular}[c]{@{}l@{}}0.5\\ 0.5\\ 0.5\\ 0.5\\ 0.3\end{tabular}                       \\\hline
Quiet street      & 70.2        &
\begin{tabular}[c]{@{}l@{}}road traffic\\ bird\end{tabular}       &
\begin{tabular}[c]{@{}l@{}}road traffic \\ voices\\ street noises\\ foot step\\ bird\\ \begin{tabular}[c]{@{}l@{}}construction \\ \hfill site noise \end{tabular}\\ door house\\ door car\end{tabular} &
\begin{tabular}[c]{@{}l@{}}1.0\\ 0.7\\ 0.7\\ 0.5\\ 0.2\\ \begin{tabular}[c]{@{}l@{}} 0.1\\~\end{tabular}\\  0.2\\ 0.2\end{tabular}
\\ \hline
Noisy street      & 73.5        & road traffic                                                      & \begin{tabular}[c]{@{}l@{}}traffic\\ foot step\\ voice\\ street noise\\ bell\\ bird\\ car horn\\ car's door\\ siren\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}9.0 \\ 0.5\\ 0.6\\ 0.4\\ 0.1\\ 0.2\\ 0.3\\ 0.2\\ 0.1\end{tabular} \\ \hline
Very noisy street & 76.0        & road traffic                                                      & \begin{tabular}[c]{@{}l@{}}traffic\\ voice\\ siren\\ car horn\\ bird\\ foot step\\ car's door\\ street noise\end{tabular}                          & \begin{tabular}[c]{@{}l@{}}40\\ 0.3\\ 0.2\\ 0.3\\ 0.2\\ 0.3\\ 0.2\\ 0.3\end{tabular}
\end{tabular}
\caption{Sound level and description of the most recurrent sound classes (backgrounds and events) present in the sound environments}
\label{tab:obsScene}
\end{table*}

As the calibration audio file was not available, the sound level cannot be considered as absolute but the relative difference between each sound environment can be considered. The road traffic, the voice and the bird components are the most discriminant classes of the sound environment: with exclusively road traffic, the \textit{noisy} and \textit{very noisy} environments sound environments are different to the \textit{park} and \textit{quiet street} where the voice and the bird's whistles are predominant. Furthermore, multiple sound classes can be heard whatever the urban sound environments : dog barks, church bell ringing, car horn, sirens \dots with a density that can be very low (< 0.1/min). Finally, a lot of brief sounds with unknown origin can be listened in almost all the files. Consequently, all these indefinite sounds are annotated in one unique class sound, \textit{street noise}. From this information (density, level, sound classes present in each sound environment), it is now possible to compose urban sound mixtures.

\section{Simulation of realistic auditory urban scenes} \label{sec:simulation}
\subsection{Presentation of the web-simulator \textit{simScene}}
\textit{simScene} \cite{rossignol_simscene:_2015} is a simulator that creates sound mixtures in a .wav format by superposing audio samples that come from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds) that are considered as salient belong to the \textit{event} category whereas all the sound that are of long duration and whose acoustic properties do not vary with respect to time belong to the \textit{background} category. Inside each category, the sound samples are then grouped in sound classes (bird, car, foot steps \dots). The software enables the user to control for each sound class some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots. Each parameter is completed with a standard deviation that may brings some random behavior between the scenes. \\

\begin{figure}[h]
\centering
\includegraphics[width = .7\linewidth]{./pictures/spectrogramme_abstract_sceneSimpleKlaxonFixe_EN.pdf}
\caption{Example of a scene composed of three sound classes}
\label{fig:exampleSimScene}
\end{figure}

The sound mixtures can be created following two modes: the \textit{abstract} mode allows to create mixtures from the specified parameters while the \textit{replicate} mode reproduces an existing scene by following as much as possible its annotation file. With the audio file of the global sound mixture, each sound class can be resumed in a separate audio file, which allows to know their exact contribution in the scene.

\subsection{Creation of the urban sound database}
Based on the annotations of the 74 real scenes, a sound database is built using the \textit{replicate} mode of \textit{simScene} with sounds found online on the \textit{freesound} project and with the help of the urbanSound8k database \cite{salamon_dataset_nodate}. This database consists in more than 8000 files with a 4 seconds or less duration, found on the web site freesound.org too, and classified in 10 sound classes : ventilation, car horn, children playing, dog barking, bell, engine idling, gunshot, jackhammer, siren and street music. All the audio files have been sorted and those who presented the best signal to noise ratio have been selected.\\

In addition, as road traffic is a prime audio source in an urban sound environment, it seemed interesting to have, in the database, a part composed of well-recorded car sounds. As a result, passages of 4 different cars (Renault Scenic, Clio, Megane and Dacia Sandero) were recorded on the Ifsttar-Nantes runway at different speeds and gear ratios for steady, acceleration, braking (Table~\ref{tab:resum_car_audio}) and stopped phases. Among the 108 recordings planned, those for the acceleration and braking phase for the Renault Senic could not be done whereas some recordings have been made twice for the other vehicles. In all, 103 car passages have been recorded. \\
